---
title: "learning_corale_IBD_rds_anovoa"
author: "Alfredo_Novoa"
date: "5/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Notes
- The only note here is whenever I restore a `lock.file()` if I have more recent versions of a library it suggests I restart my R session to run the older version. BUT once I restart it doesn't have any of the libraries loaded (even though I saved the workspace image) and requires I run renv again. I've tried running `renv::load(project = getwd())` and it says its already synced. but the packages say NOT INSTALLED in the package library.

```{r}
###############################################################
#DATA IMPORT
rm(list = ls())
getwd()
```


```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(evaluate)
library(Hmisc)
library(tableone)
library(labelled)
library(rstatix)
library(ggpubr)
library(table1)
library(hrbrthemes)
```


```{r}
#sourcing an R script to import and set up the data
#source("../code/CORALEIBD_R_2021-03-31_1200.r", local = FALSE)
#note the local version of this file defines the 'data' in the correct folder see below
data <- read_rds("../data/deid_ibdc2_27_apr.rds")

#interesting challenge was that the original script was more thorough than what I need and was given due to HIPAA guidelines
#errors in code make the source command stop. 
#using different package to push through the errors 
replay(evaluate(file("../code/copyCORALEIBD_R_2021-03-31_1200.r")))
#boss said not to use this file cuz the factor variables make it tricky
#check id_hcw_ibd in the updated one
#this creates more variables than I need which is okay. 
#ultimately the most useful stuff are the variable descriptions. 


#################################################################
#VARIABLE SETUP

#checking errors for file
evalres <- evaluate(file("../code/copyIBD_R_2021-03-31_1200.r"))
ierr <- grepl("error|warning", sapply(evalres, class))
replay(evalres[ierr]) #didnt affect key variables in the data
#errors I saw only applied to variables that were missing from my data which is perfect 

#importing and adding variable data from dictionary
dict_file <- read_csv("../data/CORALEIBD_DataDictionary_2021-03-10 (1).csv", col_names = TRUE)
dictionary_names <- dict_file$`Variable / Field Name`
dictionary_labels <- dict_file$`Field Label`
#levels of each variable is tricky. need to come back to it.

#wasn't given unique IDs for privacy reasons? 
#data is raw export from redcap so sort should stay the same 
data <- data %>% mutate(id_nrow = row_number())
glimpse(data) #checking if labels and levels it worked
levels(data$ethnicity)
labels(data$srv_household_autoimmune.factor)
###################################################################
```
I can look at 2_table_1 updated variables and append relevant vars to explore the GI symptoms after dose 1|&2 

two types of questions,
- local injection site reactions
- systemic reactions
  -- within these are the GI questions (Nausea, vomiting, diarrhea )
  
- IBD Disease types (umbrella term)
  - Crohns Disease (CD)
  - Ulcerative Colitis (UC/IC)
  
any figures should be saved as PNG.s and added to box just in case.

```{r}
##############################################################
#getting to know the dataset
typeof(data)
#str(data)
data_names <- as.tibble(names(data))

data %>% group_by(covid19_vaccine_symptoms_after_dose_1_complete) %>%
  count() %>% ungroup() #1771 complete surveys
data %>% group_by(covid19_vaccine_symptoms_after_dose_2_complete) %>%
  count() %>% ungroup() #1438 complete surveys
data %>% group_by(covid19_vaccine_symptoms_after_dose_1_complete,covid19_vaccine_symptoms_after_dose_2_complete) %>%
  count() %>% ungroup() 
#1331 participants completed both surveys
#440 completed the first dose survey
#107 complete the second dose

```


```{r}
symp_after_by_vaxtype <- vaxd_data %>% filter(is.na(vaccine_type)) %>%
  group_by(vaccine_type,
                       any_1st_dose_sx,
                       any_1st_dose_sx_v2) %>% 
  summarise(n_per_group=n(), ) %>% ungroup() 
#there are 386 who completed at least 1 questionnaire and didnt report vax type
#295 still reported some sort of new symptom after receiving a vax though.


##################################
#Other checks
vaxd_data %>% filter(!is.na(vaccine_type)) %>% group_by(id_hcw_ibd) %>% summarise(n_hcw_ibd =n()) %>% ungroup
#group 1 is IBD since the HCW group has all pfizer vaccines


###########################################################
#zeroing in on the GI symptoms
typeof(vaxd_data$post_vaccine_symptom___7)

#symptom questions
#levels(data$any_1st_dose_sx.factor)=c("Yes","No","Not sure")
#levels(data$any_1st_dose_sx_v2.factor)=c("Yes","No","Not sure")
#data$any_1st_dose_sx.factor = factor(data$any_1st_dose_sx,levels=c("1","0","2"))
#data$any_1st_dose_sx_v2.factor = factor(data$any_1st_dose_sx_v2,levels=c("1","0","2"))

vaxd_data %>% filter(!is.na(vaccine_type)) %>% 
  group_by(any_1st_dose_sx) %>% 
  summarise(n_per_group=n(), ) %>% 
  ungroup #633 had new symptoms 

vaxd_data %>% filter(!is.na(vaccine_type)) %>% 
  group_by(any_1st_dose_sx_v2) %>% 
  summarise(n_per_group=n(), ) %>% 
  ungroup #719 had new symptoms
#total new symptoms 1352
#unique=962

```

Variable Creation

```{r}
#vaxd_data <- vaxd_data %>% mutate(
 # post_vax_symptom_7_int = as.integer(post_vaccine_symptom___7),
  #nausea_timing_int = as.integer(nausea_timing),
  #vomiting_timivomiting_timing, diarrhea_timing, constipation_timing,
    #                       gerd_timing, abdo_timing)


#### Timing Levels after dose 1
 #1, Persisted for < 1 hour | 
 # 2, Persisted for < 1 day |
 # 3, Persisted for 1-2 days |
 # 4, Persisted up to 1 week |
 # 5, Persisted for >1 week |
 # 0, Never

#### Severity Levels
# 0, None |
# 1, MILD |
# 2, MODERATE |
# 3, SEVERE |
# 4, EXTREME

vaxd_data %>% filter(any_1st_dose_sx == 1 | 
           any_1st_dose_sx_v2 == 1) %>% group_by(post_vaccine_symptom___7) %>% 
  summarise(
    n_symptafterdose1 = n())
#74 people experienced gi symptoms after dose 1
#n=1257
vaxd_data %>% filter(any_1st_dose_sx == 1 | 
           any_1st_dose_sx_v2 == 1) %>% group_by(post_vaccine_symptom_v2___7) %>% 
  summarise(
    n_symptafterdose2 = n())
#184 people experienced symptoms after dose 2

###################################################################
#creating numeric variables
vaxd_data <- vaxd_data %>% mutate(across(starts_with("postvax1_local"), as.numeric),
    across(starts_with("post_vaccine_symptom_"), ~as.logical(as.numeric(.x))),
    across(contains("timing"), as.numeric),
    across(contains("sev"), as.numeric),
    across(one_of("change_in_hearing", "change_in_hearing_v2", "change_in_vision", "change_in_vision_v2"), as.numeric))

typeof(vaxd_data$nausea_timing)#type double/numeric

#vaccine type responses
#1, BNT162 (Pfizer/BioNtech) | 2, mRNA-1273 (Moderna/NIH) | 
#3, AZD12222 (AstraZeneca) | 6, JNJ-78436725 (Johnson & Johnson) |
#4, Other | 5, Not sure 
#use the orignal. nausea_sev, vomiting_sev, diarrhea_sev, constipation_sev,        stomach_burning_sev, abdominal_pain_sev,

###############################################################
#some people responded not sure to questions with branch logic so we need to resolve the "NOT SURE" Responses
vaxd_data <- vaxd_data %>%
  mutate(
    survey1 = !is.na(any_1st_dose_sx), # did they answer the first question,
    survey2 = !is.na(any_1st_dose_sx_v2), # did they answer the first question,
    age = as.numeric(age),
    any_symptom_v1 = any_1st_dose_sx,
    any_symptom_v1 = if_else(any_1st_dose_sx == 2 & all(
      post_vaccine_symptom___1 == FALSE,
      post_vaccine_symptom___2 == FALSE,
      post_vaccine_symptom___3 == FALSE,
      post_vaccine_symptom___4 == FALSE,
      post_vaccine_symptom___5 == FALSE,
      post_vaccine_symptom___6 == FALSE,
      post_vaccine_symptom___7 == FALSE,
      post_vaccine_symptom___8 == FALSE,
      post_vaccine_symptom___9 == FALSE,
      post_vaccine_symptom___10 == FALSE,
      post_vaccine_symptom___11 == FALSE,
      post_vaccine_symptom___12 == FALSE,
      postvax1_local_pain == 0,
      postvax1_local_redness == 0,
      postvax1_local_swelling == 0), "No", as.character(any_1st_dose_sx)),
    any_symptom_v1 = if_else(any_1st_dose_sx == 2 & !all(
      post_vaccine_symptom___1 == FALSE,
      post_vaccine_symptom___2 == FALSE,
      post_vaccine_symptom___3 == FALSE,
      post_vaccine_symptom___4 == FALSE,
      post_vaccine_symptom___5 == FALSE,
      post_vaccine_symptom___6 == FALSE,
      post_vaccine_symptom___7 == FALSE,
      post_vaccine_symptom___8 == FALSE,
      post_vaccine_symptom___9 == FALSE,
      post_vaccine_symptom___10 == FALSE,
      post_vaccine_symptom___11 == FALSE,
      post_vaccine_symptom___12 == FALSE,
      postvax1_local_pain == 0,
      postvax1_local_redness == 0,
      postvax1_local_swelling == 0), "Yes", as.character(any_1st_dose_sx)),
    any_symptom_v2 = any_1st_dose_sx_v2,
    any_symptom_v2 = if_else(any_1st_dose_sx_v2 == 2 & all(
      post_vaccine_symptom_v2___1 == FALSE,
      post_vaccine_symptom_v2___2 == FALSE,
      post_vaccine_symptom_v2___3 == FALSE,
      post_vaccine_symptom_v2___4 == FALSE,
      post_vaccine_symptom_v2___5 == FALSE,
      post_vaccine_symptom_v2___6 == FALSE,
      post_vaccine_symptom_v2___7 == FALSE,
      post_vaccine_symptom_v2___8 == FALSE,
      post_vaccine_symptom_v2___9 == FALSE,
      post_vaccine_symptom_v2___10 == FALSE,
      post_vaccine_symptom_v2___11 == FALSE,
      post_vaccine_symptom_v2___12 == FALSE,
      postvax1_local_pain_v2 == 0,
      postvax1_local_redness_v2 == 0,
      postvax1_local_swelling_v2 == 0), "No", as.character(any_1st_dose_sx_v2)),
    any_symptom_v2 = if_else(any_1st_dose_sx_v2 == 2 & !all(
      post_vaccine_symptom_v2___1 == FALSE,
      post_vaccine_symptom_v2___2 == FALSE,
      post_vaccine_symptom_v2___3 == FALSE,
      post_vaccine_symptom_v2___4 == FALSE,
      post_vaccine_symptom_v2___5 == FALSE,
      post_vaccine_symptom_v2___6 == FALSE,
      post_vaccine_symptom_v2___7 == FALSE,
      post_vaccine_symptom_v2___8 == FALSE,
      post_vaccine_symptom_v2___9 == FALSE,
      post_vaccine_symptom_v2___10 == FALSE,
      post_vaccine_symptom_v2___11 == FALSE,
      post_vaccine_symptom_v2___12 == FALSE,
      postvax1_local_pain_v2 == 0,
      postvax1_local_redness_v2 == 0,
      postvax1_local_swelling_v2 == 0
      ), "Yes", as.character(any_1st_dose_sx_v2))) %>%
  mutate(any_symptom_v1 = case_when(any_symptom_v1 == "Yes" ~ TRUE,
                               any_symptom_v1 == "No" ~ FALSE, 
                               TRUE ~ NA),
         any_symptom_v2 = case_when(any_symptom_v2 == "Yes" ~ TRUE,
                               any_symptom_v2 == "No" ~ FALSE, 
                               TRUE ~ NA),
         no_symptom_v1 = !(any_symptom_v1),
         no_symptom_v2 = !(any_symptom_v2))
#this makes 7 new variables mostly to check the rest of responses for accuracy to certain questions
#to be safe maybe we should select people who report any symptoms after each vaccine.

###################################################################
###############################################

#creating numeric variables
vaxd_data <- vaxd_data %>% mutate(
    across(starts_with("post_vaccine_symptom_"), ~as.logical(as.numeric(.x))),
    across(contains("timing"), as.numeric),
    across(contains("sev"), as.numeric))



############################################################################
####################################################################
#COMPUTING ANALYSIS VARIABLES
#need to group by and count TRUE sympt_7 


#will likely need to run comparisons with HCW data before filtering to check power
vaxd_data <- vaxd_data %>% rowwise() %>% 
              mutate(gi_timing_avg = mean(c(nausea_timing,
                           vomiting_timing, diarrhea_timing,
                           constipation_timing, gerd_timing,
                           abdo_timing), na.rm=TRUE),
                     gi_sev_avg = mean(c(nausea_sev, 
                           vomiting_sev, diarrhea_sev,
                           constipation_sev, stomach_burning_sev, 
                           abdominal_pain_sev),na.rm= TRUE),
                     gi_timing_avg_v2 = mean(c(nausea_timing_v2,
                           vomiting_timing_v2, diarrhea_timing_v2,
                           constipation_timing_v2, gerd_timing_v2,
                           abdo_timing_v2), na.rm=TRUE),
                     gi_sev_avg_v2 = mean(c(nausea_sev_v2,
                            vomiting_sev_v2, diarrhea_sev_v2,
                           constipation_sev_v2, stomach_burning_sev_v2, 
                           abdominal_pain_sev_v2),na.rm= TRUE))


vaxd_data%>% select(nausea_timing,vomiting_timing, diarrhea_timing,
                    constipation_timing,gerd_timing,
                    abdo_timing, gi_timing_avg, 
                    gi_sev_avg) %>% str()


vaxd_data %>% filter(any_1st_dose_sx == 1 | 
           any_1st_dose_sx_v2 == 1) %>% 
    group_by(id_hcw_ibd, post_vaccine_symptom___7) %>% 
    summarise(
      n_symptafterdose1 = n())
#74 people experienced gi symptoms
#n=1257
vaxd_data %>% filter(any_1st_dose_sx == 1 | 
           any_1st_dose_sx_v2 == 1) %>% group_by(
             id_hcw_ibd, post_vaccine_symptom_v2___7) %>% 
  summarise(
    n_symptafterdose2 = n())

vaxd_data %>% group_by(id_hcw_ibd) %>% 
  summarise(n_per_group=n(), 
            avg_gi_timing = mean(gi_timing_avg, na.rm=T),
            sd_gi_timing = sd(gi_timing_avg, na.rm = T),
            n_gi_timing = sum(!is.na(gi_timing_avg)),
            n_gi_sev = sum(!is.na(gi_sev_avg)),
            avg_gi_sev = mean(gi_sev_avg, na.rm = T),
            sd_gi_sev = sd(gi_sev_avg, na.rm = T)) %>% 
  ungroup
#This checks out. there is not a ton of variation but we can work with this
vaxd_data %>% filter(!is.na(vaccine_type) | !is.na(vaccine_type_v2)) %>% 
  group_by(id_hcw_ibd) %>% 
  summarise(n_per_group=n(), 
            avg_gi_timing = mean(gi_timing_avg, na.rm=T),
            sd_gi_timing = sd(gi_timing_avg, na.rm = T),
            n_gi_timing = sum(!is.na(gi_timing_avg)),
            n_gi_sev = sum(!is.na(gi_sev_avg)),
            avg_gi_sev = mean(gi_sev_avg, na.rm = T),
            sd_gi_sev = sd(gi_sev_avg, na.rm = T),
            n_any_yes_gi =sum(!is.na(post_vaccine_symptom___7))) %>% 
  ungroup
label(vaxd_data$post_vaccine_symptom___7)

vaxd_data %>% filter(!is.na(vaccine_type) | !is.na(vaccine_type_v2)) %>% 
  group_by(id_hcw_ibd, vaccine_type) %>% 
  summarise(n_per_group=n(), 
            avg_gi_timing = mean(gi_timing_avg, na.rm=T),
            sd_gi_timing = sd(gi_timing_avg, na.rm = T),
            n_gi_timing = sum(!is.na(gi_timing_avg)),
            n_gi_sev = sum(!is.na(gi_sev_avg)),
            avg_gi_sev = mean(gi_sev_avg, na.rm = T),
            sd_gi_sev = sd(gi_sev_avg, na.rm = T)) %>% 
  ungroup
#this anova will be very interesting to see
#should check that vaccine_type is consistently responded after each survey


vaxd_data_2 <- vaxd_data %>% filter(!is.na(vaccine_type) | !is.na(vaccine_type_v2))

vaxd_data %>% group_by(id_hcw_ibd,) %>% 
  summarise(n_per_group=n()) %>% ungroup()

##################################################
###############
#for table1 descriptive reporting
#Severity factors dose 1 & 2

vaxd_data <- vaxd_data %>%  mutate(
  nausea_sev_factor =
    factor(nausea_sev,levels=c("0","1","2","3","4")), 
  vomiting_sev_factor = 
    factor(vomiting_sev,levels=c("0","1","2","3","4")),
  diarrhea_sev_factor =
    factor(diarrhea_sev,levels=c("0","1","2","3","4")),
  constipation_sev_factor =
    factor(constipation_sev,levels=c("0","1","2","3","4")),
  stomach_burning_sev_factor =
    factor(stomach_burning_sev,levels=c("0","1","2","3","4")),
  abdominal_pain_sev_factor =
    factor(abdominal_pain_sev,levels=c("0","1","2","3","4")), #end dose 1
   nausea_sev_v2_factor =
    factor(nausea_sev_v2,levels=c("0","1","2","3","4")),
  vomiting_sev_v2_factor =
    factor(vomiting_sev_v2,levels=c("0","1","2","3","4")),
  diarrhea_sev_v2_factor =
    factor(diarrhea_sev_v2,levels=c("0","1","2","3","4")),
  constipation_sev_v2_factor =
    factor(constipation_sev_v2,levels=c("0","1","2","3","4")),
  stomach_burning_sev_v2_factor =
    factor(stomach_burning_sev_v2,levels=c("0","1","2","3","4")),
  abdominal_pain_sev_v2_factor =
    factor(abdominal_pain_sev_v2,levels=c("0","1","2","3","4")))


###########################################
#CREATING BINARY VARS for visualization and checking
#nausea_sev_binary
#vomiting_sev_binary
#diarrhea_sev_binary
#constipation_sev_binary
#stomach_burning_sev_binary
#abdominal_pain_sev_binary
#sev maxes and binaries dose 1 & 2
vaxd_data <- vaxd_data %>%  mutate(
  gi_severity_max_v1 = if_else(
    covid19_vaccine_symptoms_after_dose_1_complete != 2, NA_real_, 
          if_else(post_vaccine_symptom___7 == TRUE,
                pmax(nausea_sev, vomiting_sev, diarrhea_sev,
                 constipation_sev, stomach_burning_sev,
                 abdominal_pain_sev, na.rm=T), 0)),
  gi_severity_max_v2 = if_else(
    covid19_vaccine_symptoms_after_dose_2_complete != 2, NA_real_, 
          if_else(post_vaccine_symptom_v2___7 == TRUE,
                pmax(nausea_sev_v2, vomiting_sev_v2, diarrhea_sev_v2,
                 constipation_sev_v2, stomach_burning_sev_v2,
                 abdominal_pain_sev_v2, na.rm=T), 0)),
  nausea_sev_binary = case_when(nausea_sev > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE), #dose 1 binaries
  vomiting_sev_binary= case_when(vomiting_sev > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
  diarrhea_sev_binary= case_when(diarrhea_sev > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
  constipation_sev_binary= case_when(constipation_sev > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
    stomach_burning_sev_binary= case_when(stomach_burning_sev > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
    abdominal_pain_sev_binary=case_when(abdominal_pain_sev > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
  nausea_sev_v2_binary= case_when(nausea_sev_v2 > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE), #start dose 2 binaries
    vomiting_sev_v2_binary= case_when(vomiting_sev_v2 > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
    diarrhea_sev_v2_binary= case_when(diarrhea_sev_v2 > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
    constipation_sev_v2_binary= case_when(constipation_sev_v2 > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
    stomach_burning_sev_v2_binary= case_when(
            stomach_burning_sev_v2 > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE),
    abdominal_pain_sev_v2_binary=case_when(
            abdominal_pain_sev_v2 > 0 ~ TRUE,
             post_vaccine_symptom___7 == TRUE ~ FALSE))


vaxd_data <- vaxd_data %>%  mutate(
  gi_severity_max_v1_factor =
    factor(gi_severity_max_v1,levels=c("0","1","2","3","4")), 
  gi_severity_max_v2_factor = 
    factor(gi_severity_max_v2,levels=c("0","1","2","3","4")))

#looking at total number of symptoms per respondent
vaxd_data <- vaxd_data %>% rowwise() %>% mutate(
  gi_sev_symptom_sum = sum(c(nausea_sev_binary, vomiting_sev_binary,
                    diarrhea_sev_binary, constipation_sev_binary, 
                    stomach_burning_sev_binary,
                    abdominal_pain_sev_binary), na.rm=T),
    gi_sev_symptom_sum_v2 = sum(c(nausea_sev_v2_binary,
                  vomiting_sev_v2_binary, diarrhea_sev_v2_binary, 
                  constipation_sev_v2_binary, 
                  stomach_burning_sev_v2_binary,
                  abdominal_pain_sev_v2_binary), na.rm=T)) 


```


```{r}
#percent of gi symptoms to any symptoms? 
#need a prop table with custom n sizes based on responses to other questions in the questionnaire
#percent of gi symptoms among all respondents

prop_table1a <- vaxd_data %>% group_by(id_hcw_ibd_factor) %>% 
  summarise(data_n = round(n(), digits = 0),
        n_completed_dose1_surv = 
          round(sum(covid19_vaccine_symptoms_after_dose_1_complete==2),
                    digits = 0),
        n_completed_dose2_surv =
          round(sum(covid19_vaccine_symptoms_after_dose_2_complete==2),
                  digits = 0),
        n_any_symptoms_dose1 = round(sum(any_new_symptom_v1 == TRUE,
                                         na.rm = T), digits = 0),
        n_any_symptoms_dose2 = round(sum(any_new_symptom_v2== TRUE,
                                   na.rm = T), digits = 0),
       n_gi_sx_dose1 = round(sum(as.numeric(post_vaccine_symptom___7)),
                             digits = 0), 
        n_gi_sx_dose2 =
          round(sum(as.numeric(post_vaccine_symptom_v2___7)),
                              digits = 0),
       mean_gi_sev_symptom_n = 
          round(sum(gi_sev_symptom_sum, na.rm= TRUE)/
            sum(as.numeric(post_vaccine_symptom___7), na.rm = TRUE), digits = 2),
       mean_gi_sev_symptom_n_v2 = 
          round(sum(gi_sev_symptom_sum_v2, na.rm= TRUE)/
            sum(as.numeric(post_vaccine_symptom_v2___7), na.rm = TRUE), digits = 2),
       blank2 = NA,
       blank = NA,
       pct_new_symptoms_dose1=
            round(sum(any_new_symptom_v1 == TRUE, na.rm = T)/
              sum(covid19_vaccine_symptoms_after_dose_1_complete==2,
              na.rm = T)*100, digits = 2),
        pct_new_symptoms_dose2=
              round(sum(as.numeric(any_new_symptom_v2 == TRUE), 
              na.rm = T)/sum(
              covid19_vaccine_symptoms_after_dose_2_complete==2,
                  na.rm = T)*100, digits = 2),
        pct_gi_symptoms_dose1=
              round(sum(as.numeric(post_vaccine_symptom___7))/
              sum(covid19_vaccine_symptoms_after_dose_1_complete==2, 
                  na.rm = T)*100, digits = 2),
        pct_gi_symptoms_dose2=
              round(sum(as.numeric(post_vaccine_symptom_v2___7))/
              sum(covid19_vaccine_symptoms_after_dose_2_complete==2, 
                  na.rm = T)*100, digits = 2))



prop_table1a <- prop_table1a %>%
  labelled::set_variable_labels(
    n_completed_dose1_surv = "Count of Completed Surveys, Dose 1",
    n_completed_dose2_surv = "Count of Completed Surveys, Dose 2",
    n_any_symptoms_dose1 = "Count of ANY NEW Adverse Event after Dose 1",
     n_any_symptoms_dose2 = "Count of Any New Adverse Event after Dose 2",
    n_gi_sx_dose1 = "Count of People with New GI Symptoms after Dose 1",
    n_gi_sx_dose2 = "Count of People with New GI Symptoms after Dose ",
     mean_gi_sev_symptom_n = 
      "Average number of GI symptoms for Dose 1",
    mean_gi_sev_symptom_n_v2 =
      "Average number of GI Symptoms for Dose 2",
    pct_new_symptoms_dose1=
        "Percent of People with ANY NEW Symptoms who completed Dose 1 Survey",
    pct_new_symptoms_dose2=
        "Percent of People with ANY NEW Symptoms who completed Dose 2 Survey",
    pct_gi_symptoms_dose1 =
      "Percent of People with GI Symptoms who completed Dose 1 Survey",
    pct_gi_symptoms_dose2 =
      "Percent of People with GI Symptoms who completed Dose 2 Survey")


prop_table1b <- pivot_longer(prop_table1a, cols=2:16, names_to = "name", values_to = "value")

#replacing initial var names w descriptions
prop_table1b$name <- str_replace_all(prop_table1b$name, c(
    "n_completed_dose1_surv" =
              "n of Completed Surveys Dose 1", 
    "n_completed_dose2_surv" =
                "n of Completed Surveys Dose 2",
    "n_any_symptoms_dose1" =
                "n of ANY NEW Adverse Event after Dose 1",
    "n_any_symptoms_dose2" =
            "n of ANY NEW Adverse Event after Dose 2",
    "n_gi_sx_dose1" =
            "n of New GI Symptoms after Dose 1",
    "n_gi_sx_dose2" = 
            "n of New GI Symptoms after Dose 2",
    "mean_gi_sev_symptom_n" = 
            "Average number of GI symptoms for Dose 1",
    "mean_gi_sev_symptom_n_v2" = 
          "Average number of GI Symptoms for Dose 2",
    "blank2"= "",
    "blank"= "PERCENTAGES based on COMPLETED Surveys",
    "pct_new_symptoms_dose1"=
      "% Reporting ANY NEW Symptoms after Dose 1",
    "pct_new_symptoms_dose2"=
        "% Reporting ANY NEW Symptoms after Dose 2",
        "pct_gi_symptoms_dose1" =
        "% Reporting GI Symptoms after Dose 1",
    "pct_gi_symptoms_dose2" =
      "% Reporting GI Symptoms after Dose 2"
        )) 


prop_table1c <- prop_table1b %>%  pivot_wider(names_from =
                                  "id_hcw_ibd_factor",values_from =
                                    "value") %>% as_tibble()

prop_table1c
#top level view of data
```

Creating tables with Table 1 (see package info for details)
```{r}
#lets also break down the demographics

table1(~ sex+ 
         age +
         srv_race +
          ethnicity | id_hcw_ibd_factor, 
        data=vaxd_data, overall="Total", droplevels = FALSE)




tbl1_dta_dose1 <- vaxd_data %>% filter(covid19_vaccine_symptoms_after_dose_1_complete == 2)

table1(~ any_new_symptom_v1 +
         post_vaccine_symptom___7 +
        nausea_sev_binary +
        vomiting_sev_binary +
        diarrhea_sev_binary +
        constipation_sev_binary +
        stomach_burning_sev_binary +
        abdominal_pain_sev_binary +
         gi_severity_max_v1_factor +
         gi_timing_max_v1_factor +
         gi_sev_symptom_sum | id_hcw_ibd_factor, 
        data=tbl1_dta_dose1, overall="Total", droplevels = FALSE)


```
Also running some chi sq tests 
```{r}


vars_cat_dose1 <- c("post_vaccine_symptom___7", 
                    "nausea_sev_binary",
                    "vomiting_sev_binary", "diarrhea_sev_binary",
                    "constipation_sev_binary",
                    "stomach_burning_sev_binary", 
                    "abdominal_pain_sev_binary" )

vars_dose1 <- c("post_vaccine_symptom___7", 
                "nausea_sev_binary",
                "vomiting_sev_binary", "diarrhea_sev_binary",
                "constipation_sev_binary", "stomach_burning_sev_binary",
                "abdominal_pain_sev_binary", "gi_sev_symptom_sum")

dose1_n_by_sympt <- CreateTableOne(vars = vars_dose1,
                                   data = tbl1_dta_dose1,
                                   strata = "id_hcw_ibd_factor",
                                   factorVars = vars_cat_dose1)

dose1_n_by_sympt

chisq.test(tbl1_dta_dose1$id_hcw_ibd_factor,tbl1_dta_dose1$post_vaccine_symptom___7)
```


A few visualizations 
```{r}
#################
#stacked barchart with table maxes 
stackedchart_tablemaxes <- tbl1_dta_dose1 %>% group_by(id_hcw_ibd_factor) %>% 
  summarise( 
     max_mild_severity = 
        round(sum(gi_severity_max_v1_factor =="MILD", na.rm = TRUE),
              digits = 1),
      max_moderate_severity = 
        round(sum(gi_severity_max_v1_factor =="MODERATE", na.rm = TRUE),
              digits = 1), 
     max_severe_severity = 
        round(sum(gi_severity_max_v1_factor =="SEVERE", na.rm = TRUE),
              digits = 1), 
     max_extreme_severity = 
        round(sum(gi_severity_max_v1_factor =="EXTREME", na.rm = TRUE),
              digits = 1))
stackedchart_tablemaxesb <- pivot_longer(stackedchart_tablemaxes, cols=2:5, names_to = "name", values_to = "value")
#levels(vaxd_data$gi_severity_max_v1_factor)=c("None","MILD","MODERATE","SEVERE","EXTREME")


stackedchart_tablemaxesc <- stackedchart_tablemaxesb %>%  pivot_wider(names_from =
                                  "id_hcw_ibd_factor", values_from =
                                    "value") %>% as_tibble() 
stackedchart_tablemaxesd <-column_to_rownames(stackedchart_tablemaxesc,var = "name")
stackedchart_tablemaxesd <- as.matrix(stackedchart_tablemaxesd)

#outputs

stackedchart_tablemaxesd

barplot(stackedchart_tablemaxesd, 
        col=palette.colors(n= 4 , palette= "Pastel 2") , 
        border="white", 
        space=0.04, 
        font.axis=2, 
        xlab="group",
        ylab = "",
        axisnames = TRUE,
        legend.text = TRUE,
        args.legend = list(title= "Max GI Symptom Severity Dose 1", x = "right", cex = .7 ))

# Get the stacked barplot
#groups needs to be each column
#var values need to be rows

#table for stacked proportions of Max values
stacked_prop_maxes <- round(prop.table(stackedchart_tablemaxesd,
                                 margin = 2), digits = 3)

stacked_prop_maxes
barplot(stacked_prop_maxes, 
        col=palette.colors(n= 4 , palette= "Pastel 2") , 
        border="white", 
        space=0.04, 
        font.axis=2, 
        xlab="group",
        ylab = "",
        axisnames = TRUE,
        legend.text = TRUE,
        args.legend = list(title= "Max GI Symptom Severity Dose 1", x = "right", cex = .7 ))

```





Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
