---
title: "Make Up problem set for ec"
author: "Alfredo Novoa"
date: "December 7, 2018"
urlcolor: blue
output: 
  pdf_document:
    toc: true 
    toc_depth: 1
    df_print: tibble 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Assignment Overview: From Professor Jaquette's Emails

*"I want you to think of a __user-written function or loop__ to do something useful/interesting with data [can be data from this class or data from some other source]. Has to be more substantial than the functions/loops I had you work on for homework, but not too far beyond that.*
 
*Then I want you to develop a beta version of that function or loop. And then you will turn in two functions or two loops or one function and one loop.*

*Due within 24 hours.*


##Step 1: Select and Download Data

**for this assignment **
Data source: College Scorecard Data: https://collegescorecard.ed.gov/data/

Most Recent Cohort Import URL:https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-All-Data-Elements.csv  
Plus the data dictionary for the most recent cohort data 

*May not work out the gate so we need to make sure.* 
Objective: Add variable labels to the dataset (checked with glimpse()).
Based on the Bonus Question for Problem set 8


```{r }

rm(list= ls())

library(readr)
library(tidyverse)
library(haven)
library(labelled)

## -----------------------------------------------------------------------------
## -----------------------------------------------------------------------------
#downloading scorecard data without overwriting directories
## -----------------------------------------------------------------------------
## -----------------------------------------------------------------------------


## CHOOSE OUTPUT DIRECTORY (DEFAULT == '.', which is current directory)
getwd()
#linux pc "/home/freddy/Documents/UCLA/rclass_fall18/rclass_directory/problemsets/Make Up Assignment Brainstorm"
out_dir = '.'
## default values for our functions 
primary_data=TRUE
scorecard_data = TRUE
dictionary = TRUE
## overwrite already downloaded files
overwrite = FALSE
## message print with adjusted layout for formatting purposes to make it easy to see when a new step has begun
mess <- function(to_screen) {
  message(rep('-',80))#creates an 80 char dash format
  message(to_screen)
  message(rep('-',80))
}


## create folders if they don't exist
## creating a subdirectory function
mess('Creating directories for downloaded files')
make_dir <- function(opt, dir_name) {
  if (opt & dir.exists(dir_name)) {
    message(paste0('Already have directory: ', dir_name))
  } else if (opt & !dir.exists(dir_name)) {
    message(paste0('Creating directory: ', dir_name))
    dir.create(dir_name)
  }
}
## initialize potential file paths
data_dir <- file.path(out_dir, 'data');typeof(data_dir) #define data folder path
scorecard_dir <- file.path(out_dir, 'scorecard_dir') #define scorecard folder path
dictionary_dir <- file.path(out_dir,'dictionary') #define dictionary folder path
## checking and creating
typeof(data_dir)
make_dir(TRUE, out_dir)#testing
make_dir(primary_data, data_dir)
make_dir(scorecard_data, scorecard_dir)
#this doesn't show up but it should work once I call the file

make_dir(dictionary, dictionary_dir)



mess('Time to download the files!')

get_file <- function(opt, dir_name, url, file, suffix, overwrite) {
  if (opt) {
    dest <- file.path(dir_name, paste0(file, suffix))
    if (file.exists(dest) & !overwrite) {
      message(paste0('Already have file: ', dest))
      return(0)
    } else {
      download.file(paste0(url, file, suffix), dest)#notice file name cat()s , good for batch url exports 
      Sys.sleep(1)
      return(1)
    }
  }
}
#note: the URL structure in this function depends on the data host to make data available in the same folder

scdata_file_url <- "https://ed-public-download.app.cloud.gov/downloads/"
#not included because of function str:'Most-Recent-Cohorts-All-Data-Elements.csv'
dict_file_url <- "https://collegescorecard.ed.gov/assets/" 
#not included because of function str: 'CollegeScorecardDataDictionary.xlsx'

## data
sc_recent_cohort <- get_file(scorecard_data, scorecard_dir, scdata_file_url, 'Most-Recent-Cohorts-All-Data-Elements', '.csv', overwrite)

sc_cohort_dictionary <- get_file(dictionary, dictionary_dir, dict_file_url, 'CollegeScorecardDataDictionary', '.xlsx', overwrite)

mess("Now lets Import a subsample of the the Dataset file")
#Review documentation before reading in data! 
#sc_mrcohort <- read_csv("./scorecard_dir/Most-Recent-Cohorts-All-Data-Elements.csv") #this didn't work
sc_mrcohort2 <- as.tibble(read.csv("./scorecard_dir/Most-Recent-Cohorts-All-Data-Elements.csv", 
                        header = TRUE, fill = TRUE, comment.char = ""))

names(sc_mrcohort2) <- tolower(names(sc_mrcohort2))
sc_mrcohort2 <- sc_mrcohort2 %>%
  select(opeid6, opeid, instnm, city, stabbr, zip, region, latitude,longitude, ccbasic, adm_rate, adm_rate_all,
         pcip01, pcip03, pcip04, pcip05, pcip09, pcip10, pcip11, pcip12, pcip13, pcip14, pcip15, pcip16, pcip19,
         pcip22, pcip23, pcip24, pcip25, pcip26, pcip27, pcip29, pcip30, pcip31, pcip38, pcip39, pcip40, pcip41,
         pcip42, pcip43, pcip44, pcip45, pcip46, pcip47, pcip48, pcip49, pcip50, pcip51, pcip52)

#Change names to lowercase  
names(sc_mrcohort2) 


head(sc_mrcohort) 

#next steps
#import the dictionary file.
#make everything lowercase 
#use the variable names column to match the name to the var description.
#try looping and using 'else if' to check if the variable name in the dictionary sheet is in the current dataset
#if yes, then append the variable description to the var_label
#ran out of time but I hit


```
Sorry Profe, I ran out of time to finish the dictionary variable names import. However, I believe I've met the criteria for the bonus assignment, I hope that's okay.